### 2021-04-09 ch03_esmvaltool/esmvaltool.tex
Capitalized manual references of figures, tables, chapters, sections and versions

The \ac{ESMValTool} is developed open-source on GitHub, a web service that
provides hosting for version control with \emph{git}
(\url{https://github.com/ESMValGroup}). It is released under the Apache
License, version 2.0. The source code of the latest released version of the
\ac{ESMValCore} package, which includes the core functionalities of the
\ac{ESMValTool}, is publicly available at Zenodo \autocite{Andela2020a}.
Similarly, the source code of the latest released version of the
\ac{ESMValTool} package, which includes the recipe and diagnostic scripts that
can for example be used to reproduce scientific assessments, is also available
at Zenodo \autocite{Andela2020}.

@@

\begin{table}[t]
  \centering
  \begin{tabulary}{\columnwidth}{L L L}
    \toprule
    Type & Name & Main reference \\
    \midrule
    Recipe & \emph{bock20jgr} (figures 8 and 10) & \autocite{Bock2020} \\
    & \emph{cox18nature} & \autocite{Cox2018} \\
    & \emph{ecs} & \autocite{Gregory2004} \\
    & \emph{flato13ipcc} (figure 9.42) & \autocite{Flato2013} \\
    & \emph{meehl20sciadv} & \autocite{Meehl2020} \\
    & \emph{schlund20esd} & \autocite{Schlund2020a} \\
    & \emph{schlund20jgr\_gpp\_abs\_rcp85} & \autocite{Schlund2020} \\
    & \emph{schlund20jgr\_gpp\_change\_1pct} & \autocite{Schlund2020} \\
    & \emph{schlund20jgr\_gpp\_change\_rcp85} & \autocite{Schlund2020} \\
    & \emph{tcr} & \autocite{Gregory2008} \\
    \midrule
    \acs{CMOR}izer & CRU & \autocite{Harris2014} \\
    & CT2019 & \autocite{Jacobson2020} \\
    & GCP & \autocite{Friedlingstein2020} \\
    & HWSD & \autocite{Wieder2014} \\
    & JMA-TRANSCOM & \autocite{Maki2010} \\
    & LAI3g & \autocite{Zhu2013} \\
    & LandFlux-EVAL & \autocite{Mueller2013} \\
    & MLS-Aura & \autocite{Read2015} \\
    & FLUXNET-MTE & \autocite{Jung2011} \\
    & NDP & \autocite{Gibbs2006} \\
    & Scripps-\acs{CO2} & \autocite{Keeling2005} \\
    \bottomrule
  \end{tabulary}
  \caption[
    Summary of new recipes and \acs{CMOR}ization scripts for observational data
    contributed to the \acf{ESMValTool} by the author of this thesis.
  ]{
    Summary of new recipes and \acs{CMOR}ization scripts for observational data
    contributed to the \acf{ESMValTool} by the author of this thesis. The
    corresponding file names of the recipes in the \acs{ESMValTool} repository
    are given by \emph{recipe\_(name\_in\_table).yml}.
  }
  \label{tab:03:changes_to_esmvaltool}
\end{table}

Due to an exhaustive overhaul of the \ac{ESMValTool} version 2, older recipes
from version 1 \autocite{Eyring2016a} cannot be used directly in the latest
release. Porting recipes from the old version to the current version requires
substantial changes on the recipes themselves, but also on the corresponding
diagnostic scripts. As of \TheMonth{}, the author of this thesis has ported two
recipes originally written by other authors to the new version of
\ac{ESMValTool}: \emph{recipe\_anav13jclim.yml} and
\emph{recipe\_wenzel16jclim.yml}. \emph{recipe\_anav13jclim.yml} reproduces the
analysis of \textcite{Anav2013}, who evaluate carbon cycle--related variables
for \acp{ESM} of the \acs{CMIP}5 ensemble. This includes climatologies, trends,
and variabilities of important quantities of the terrestrial and oceanic carbon
cycle like carbon fluxes, carbon reservoir sizes, and vegetation distributions.
\emph{recipe\_wenzel16jclim.yml} includes the \ac{MDER} analysis performed by
\textcite{Wenzel2016a} to constrain future changes in the position of the
austral jet stream. Due to its flexible code, the corresponding diagnostic
scripts can be used to apply the \ac{MDER} method to arbitrary target variables
and predictors. Both recipes include analyses that contributed ideas and
methods for the study presented in \cref{ch:06:paper_gpp}, in which the
\ac{MDER} method is expanded with \iac{ML} approach and applied to a carbon
cycle--related target variable (\ac{GPP}).

@@

An important class of recipes in the \ac{ESMValTool} covers international
climate assessments. One such example is \emph{recipe\_flato13ipcc.yml}, which
reproduces large parts of the climate model evaluation performed in chapter 9
of the \ac{IPCC}'s Fifth \acl{AR} (\acs{AR}5) \autocite{Flato2013}. For this
recipe, the two-panel figure 9.42 has been added, which shows the relationship
between the historical and pre-industrial \ac{GSAT} and \ac{ECS}
(\cref{fig:03:flato13:a}) and the relationship between \ac{TCR} and \ac{ECS}
(\cref{fig:03:flato13:b}) within the \acs{CMIP}5 ensemble.
\Cref{fig:03:flato13:a} illustrates that there is no clear connection between
\ac{ECS} and the historical or pre-industrial \ac{GSAT}, \ie{} models with high
temperatures in the historical or pre-industrial period do not necessarily
exhibit a high equilibrium warming in the future or vice versa. On the other
hand, \ac{TCR} and \ac{ECS} are well correlated within the \acs{CMIP}5 ensemble
as theoretically expected from \cref{eq:02:tcr_vs_ecs}. However, this figure
only shows a linear regression fit between the two variables and not the
expected reciprocal relation. The output of \emph{recipe\_flato13ipcc.yml}
provided useful input for the analysis of climate sensitivity in \acs{CMIP}6
performed in \textcite{Meehl2020}, which is also presented in
\cref{ch:04:papers_ecs_tcr_assessment}.

>>>>

The \ac{ESMValTool} is developed open-source on GitHub, a web service that
provides hosting for version control with \emph{git}
(\url{https://github.com/ESMValGroup}). It is released under the Apache
License, Version 2.0. The source code of the latest released version of the
\ac{ESMValCore} package, which includes the core functionalities of the
\ac{ESMValTool}, is publicly available at Zenodo \autocite{Andela2020a}.
Similarly, the source code of the latest released version of the
\ac{ESMValTool} package, which includes the recipe and diagnostic scripts that
can for example be used to reproduce scientific assessments, is also available
at Zenodo \autocite{Andela2020}.

@@

\begin{table}[t]
  \centering
  \begin{tabulary}{\columnwidth}{L L L}
    \toprule
    Type & Name & Main reference \\
    \midrule
    Recipe & \emph{bock20jgr} (Figures 8 and 10) & \autocite{Bock2020} \\
    & \emph{cox18nature} & \autocite{Cox2018} \\
    & \emph{ecs} & \autocite{Gregory2004} \\
    & \emph{flato13ipcc} (Figure 9.42) & \autocite{Flato2013} \\
    & \emph{meehl20sciadv} & \autocite{Meehl2020} \\
    & \emph{schlund20esd} & \autocite{Schlund2020a} \\
    & \emph{schlund20jgr\_gpp\_abs\_rcp85} & \autocite{Schlund2020} \\
    & \emph{schlund20jgr\_gpp\_change\_1pct} & \autocite{Schlund2020} \\
    & \emph{schlund20jgr\_gpp\_change\_rcp85} & \autocite{Schlund2020} \\
    & \emph{tcr} & \autocite{Gregory2008} \\
    \midrule
    \acs{CMOR}izer & CRU & \autocite{Harris2014} \\
    & CT2019 & \autocite{Jacobson2020} \\
    & GCP & \autocite{Friedlingstein2020} \\
    & HWSD & \autocite{Wieder2014} \\
    & JMA-TRANSCOM & \autocite{Maki2010} \\
    & LAI3g & \autocite{Zhu2013} \\
    & LandFlux-EVAL & \autocite{Mueller2013} \\
    & MLS-Aura & \autocite{Read2015} \\
    & FLUXNET-MTE & \autocite{Jung2011} \\
    & NDP & \autocite{Gibbs2006} \\
    & Scripps-\acs{CO2} & \autocite{Keeling2005} \\
    \bottomrule
  \end{tabulary}
  \caption[
    Summary of new recipes and \acs{CMOR}ization scripts for observational data
    contributed to the \acf{ESMValTool} by the author of this thesis.
  ]{
    Summary of new recipes and \acs{CMOR}ization scripts for observational data
    contributed to the \acf{ESMValTool} by the author of this thesis. The
    corresponding file names of the recipes in the \acs{ESMValTool} repository
    are given by \emph{recipe\_(name\_in\_table).yml}.
  }
  \label{tab:03:changes_to_esmvaltool}
\end{table}

Due to an exhaustive overhaul of the \ac{ESMValTool} Version 2, older recipes
from Version 1 \autocite{Eyring2016a} cannot be used directly in the latest
release. Porting recipes from the old version to the current version requires
substantial changes on the recipes themselves, but also on the corresponding
diagnostic scripts. As of \TheMonth{}, the author of this thesis has ported two
recipes originally written by other authors to the new version of
\ac{ESMValTool}: \emph{recipe\_anav13jclim.yml} and
\emph{recipe\_wenzel16jclim.yml}. \emph{recipe\_anav13jclim.yml} reproduces the
analysis of \textcite{Anav2013}, who evaluate carbon cycle--related variables
for \acp{ESM} of the \acs{CMIP}5 ensemble. This includes climatologies, trends,
and variabilities of important quantities of the terrestrial and oceanic carbon
cycle like carbon fluxes, carbon reservoir sizes, and vegetation distributions.
\emph{recipe\_wenzel16jclim.yml} includes the \ac{MDER} analysis performed by
\textcite{Wenzel2016a} to constrain future changes in the position of the
austral jet stream. Due to its flexible code, the corresponding diagnostic
scripts can be used to apply the \ac{MDER} method to arbitrary target variables
and predictors. Both recipes include analyses that contributed ideas and
methods for the study presented in \cref{ch:06:paper_gpp}, in which the
\ac{MDER} method is expanded with \iac{ML} approach and applied to a carbon
cycle--related target variable (\ac{GPP}).

@@

An important class of recipes in the \ac{ESMValTool} covers international
climate assessments. One such example is \emph{recipe\_flato13ipcc.yml}, which
reproduces large parts of the climate model evaluation performed in Chapter 9
of the \ac{IPCC}'s Fifth \acl{AR} (\acs{AR}5) \autocite{Flato2013}. For this
recipe, the Figure 9.42 of \textcite{Flato2013} has been added, which shows the
relationship between the historical and pre-industrial \ac{GSAT} and \ac{ECS}
(\cref{fig:03:flato13:a}) and the relationship between \ac{TCR} and \ac{ECS}
(\cref{fig:03:flato13:b}) within the \acs{CMIP}5 ensemble.
\Cref{fig:03:flato13:a} illustrates that there is no clear connection between
\ac{ECS} and the historical or pre-industrial \ac{GSAT}, \ie{} models with high
temperatures in the historical or pre-industrial period do not necessarily
exhibit a high equilibrium warming in the future or vice versa. On the other
hand, \ac{TCR} and \ac{ECS} are well correlated within the \acs{CMIP}5 ensemble
as theoretically expected from \cref{eq:02:tcr_vs_ecs}. However, this figure
only shows a linear regression fit between the two variables and not the
expected reciprocal relation. The output of \emph{recipe\_flato13ipcc.yml}
provided useful input for the analysis of climate sensitivity in \acs{CMIP}6
performed in \textcite{Meehl2020}, which is also presented in
\cref{ch:04:papers_ecs_tcr_assessment}.

### 2021-04-09 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Capitalized manual references of figures, tables, chapters, sections and versions

To reduce uncertainties in multi-model climate projections, a first important
step is the assessment of the desired target variables in the corresponding
climate model ensemble. This is particularly relevant when a new generation of
\acp{ESM} is published that includes considerable modifications compared to the
respective predecessor model versions. Currently, we are in a situation like
this with the \acs{CMIP}6 ensemble, for which new data is still released every
day (as of \TheMonth{}). In light of the upcoming international climate
assessment of the \ac{IPCC}'s \acs{AR}6, the evaluation of policy-relevant
metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6 models is crucial
since they provide vital information about the future climate of the planet
Earth. As the \ac{ESMValTool} is an excellent tool that allows for a quick and
robust evaluation of new \ac{CMIP} data as soon as it gets published on the
\ac{ESGF} servers (see \cref{sec:03:esmvaltool}), it is used here to assess and
analyze \ac{ECS} and \ac{TCR} of the \acs{CMIP}6 models and compare it to
corresponding results of predecessor model generations and international
climate assessments. This work is already published in two scientific papers:
\textcite{Bock2020} and \textcite{Meehl2020}. For \textcite{Bock2020}, the
author of this thesis contributed their figure 8 (a bar chart showing the
\ac{ECS} for several \acs{CMIP}3, \acs{CMIP}5, and \acs{CMIP}6 models), their
figure 10 (map plots showing cloud-related feedback parameters for the
\acs{CMIP}5 and \acs{CMIP}6 \acp{MMM}; see
\cref{fig:04:cloud_feedback_parameters}), code to create these figures with the
\ac{ESMValTool}, and text for the manuscript (in particular a section about the
evaluation of \ac{ECS}). For \textcite{Meehl2020}, the author of this thesis
contributed all figures and tables, code to produce these with the
\ac{ESMValTool}, and text for the manuscript. This
\lcnamecref{ch:04:papers_ecs_tcr_assessment} first presents the evaluation of
climate sensitivity for the \acs{CMIP}5 and \acs{CMIP}6 ensembles
(\cref{sec:04:evaluation_ecs_and_tcr}) and puts the corresponding \ac{ECS} and
\ac{TCR} values into historical context (\cref{sec:04:historical_context}).
Finally, possible reasons for the apparent increase of climate sensitivity in
the \acs{CMIP}6 models are discussed in detail
(\cref{sec:04:possible_reasons_high_ecs_cmip6}).

@@

Various improvements of the underlying physical, biological, and chemical
processes have been introduced to the \acs{CMIP}6 \acp{ESM} to represent the
coupled Earth system in more detail. Since many of these processes directly
influence the models' \ac{ECS} \autocite{Forster2020}, determining reasons for
the increased \ac{ECS} in the \acs{CMIP}6 ensemble is highly non-trivial. As
discussed in detail by \textcite{Meehl2020}, one possible reason is the
consideration of aerosol-cloud interactions. New prognostic aerosol schemes
added to some \acs{CMIP}6 models that include aerosol-cloud interactions could
have produced overly large negative radiative forcing, which then required a
stronger model response to \ac{GHG} forcing to reproduce the observed
historical temperature trend. In fact, \textcite{Meehl2020} find a weak
inter-model relationship between the aerosol forcing and \ac{ECS} for some
\acs{CMIP}6 models, with larger negative present-day aerosol forcing associated
with higher \ac{ECS}. Such relationships have also been found for previous
climate model generations \autocite{Kiehl2007, Forster2013}. However, due to
the varying aerosol forcing over time, these relations cannot be used to tune
\ac{ECS} based on the models' responses to aerosols over the different periods
of the \nth{20} century \autocite{Dittus2020}. Further aerosol-cloud
interactions that are relevant in the context of high climate sensitivity in
\acs{CMIP}6 are found for example in the \ac{HadGEM} version 3. The new aerosol
scheme in this model suppresses a strong negative cloud feedback over the
Southern Ocean that is found in the model's predecessor versions
\autocite{BodasSalcedo2019}. The Southern Ocean is one of the most pristine
aerosol regions in the world \autocite{Hamilton2014}, where aerosols
predominantly originate from natural sources like marine emissions of gaseous
dimethyl sulfide. Earlier \ac{HadGEM} versions show a large reduction in cloud
droplet size with warming over the Southern Ocean, which increases the cloud
albedo and thus produces a negative shortwave cloud feedback. On the contrary,
in \acs{HadGEM}3, this negative feedback is completely suppressed or even
positive due to a small increase of the cloud droplet size with warming, which
eventually increases \ac{ECS} due to an increased net cloud feedback
\autocite{BodasSalcedo2019}.

>>>>

To reduce uncertainties in multi-model climate projections, a first important
step is the assessment of the desired target variables in the corresponding
climate model ensemble. This is particularly relevant when a new generation of
\acp{ESM} is published that includes considerable modifications compared to the
respective predecessor model versions. Currently, we are in a situation like
this with the \acs{CMIP}6 ensemble, for which new data is still released every
day (as of \TheMonth{}). In light of the upcoming international climate
assessment of the \ac{IPCC}'s \acs{AR}6, the evaluation of policy-relevant
metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6 models is crucial
since they provide vital information about the future climate of the planet
Earth. As the \ac{ESMValTool} is an excellent tool that allows for a quick and
robust evaluation of new \ac{CMIP} data as soon as it gets published on the
\ac{ESGF} servers (see \cref{sec:03:esmvaltool}), it is used here to assess and
analyze \ac{ECS} and \ac{TCR} of the \acs{CMIP}6 models and compare it to
corresponding results of predecessor model generations and international
climate assessments. This work is already published in two scientific papers:
\textcite{Bock2020} and \textcite{Meehl2020}. For \textcite{Bock2020}, the
author of this thesis contributed their Figure 8 (a bar chart showing the
\ac{ECS} for several \acs{CMIP}3, \acs{CMIP}5, and \acs{CMIP}6 models), their
Figure 10 (map plots showing cloud-related feedback parameters for the
\acs{CMIP}5 and \acs{CMIP}6 \acp{MMM}; see
\cref{fig:04:cloud_feedback_parameters}), code to create these figures with the
\ac{ESMValTool}, and text for the manuscript (in particular a section about the
evaluation of \ac{ECS}). For \textcite{Meehl2020}, the author of this thesis
contributed all figures and tables, code to produce these with the
\ac{ESMValTool}, and text for the manuscript. This
\lcnamecref{ch:04:papers_ecs_tcr_assessment} first presents the evaluation of
climate sensitivity for the \acs{CMIP}5 and \acs{CMIP}6 ensembles
(\cref{sec:04:evaluation_ecs_and_tcr}) and puts the corresponding \ac{ECS} and
\ac{TCR} values into historical context (\cref{sec:04:historical_context}).
Finally, possible reasons for the apparent increase of climate sensitivity in
the \acs{CMIP}6 models are discussed in detail
(\cref{sec:04:possible_reasons_high_ecs_cmip6}).

@@

Various improvements of the underlying physical, biological, and chemical
processes have been introduced to the \acs{CMIP}6 \acp{ESM} to represent the
coupled Earth system in more detail. Since many of these processes directly
influence the models' \ac{ECS} \autocite{Forster2020}, determining reasons for
the increased \ac{ECS} in the \acs{CMIP}6 ensemble is highly non-trivial. As
discussed in detail by \textcite{Meehl2020}, one possible reason is the
consideration of aerosol-cloud interactions. New prognostic aerosol schemes
added to some \acs{CMIP}6 models that include aerosol-cloud interactions could
have produced overly large negative radiative forcing, which then required a
stronger model response to \ac{GHG} forcing to reproduce the observed
historical temperature trend. In fact, \textcite{Meehl2020} find a weak
inter-model relationship between the aerosol forcing and \ac{ECS} for some
\acs{CMIP}6 models, with larger negative present-day aerosol forcing associated
with higher \ac{ECS}. Such relationships have also been found for previous
climate model generations \autocite{Kiehl2007, Forster2013}. However, due to
the varying aerosol forcing over time, these relations cannot be used to tune
\ac{ECS} based on the models' responses to aerosols over the different periods
of the \nth{20} century \autocite{Dittus2020}. Further aerosol-cloud
interactions that are relevant in the context of high climate sensitivity in
\acs{CMIP}6 are found for example in the \ac{HadGEM} Version 3. The new aerosol
scheme in this model suppresses a strong negative cloud feedback over the
Southern Ocean that is found in the model's predecessor versions
\autocite{BodasSalcedo2019}. The Southern Ocean is one of the most pristine
aerosol regions in the world \autocite{Hamilton2014}, where aerosols
predominantly originate from natural sources like marine emissions of gaseous
dimethyl sulfide. Earlier \ac{HadGEM} versions show a large reduction in cloud
droplet size with warming over the Southern Ocean, which increases the cloud
albedo and thus produces a negative shortwave cloud feedback. On the contrary,
in \acs{HadGEM}3, this negative feedback is completely suppressed or even
positive due to a small increase of the cloud droplet size with warming, which
eventually increases \ac{ECS} due to an increased net cloud feedback
\autocite{BodasSalcedo2019}.
